---
title: "p8105_hw3_raj2152"
author: "Rachelle Juan"
date: "10/20/2021"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

### Load the Instacart data for `p8105.datasets`
```{r}
data("instacart")
```

The Instacart dataset has `r nrow(instacart)` observations and `r ncol(instacart)` variables. Each row or observation corresponds to a single product within an Instacart order, therefore, one order may map to several products (i.e., there is a one-to-many relationship between orders and products). The variables provide details on each order, including things like `order_id` and `product_id` as well as the `user_id` of who placed the order. The variables also include information on where each product is located, for example: `aisle` and `department`.

There are `r length(unique(pull(instacart, aisle_id)))` aisles. The aisles with the most items ordered from are: fresh vegetables, fresh fruits, packaged vegetables fruits, yogurt, and packaged cheese. 
```{r}
aisle_df = 
  instacart %>%
  count(aisle) %>% 
  arrange(desc(n)) %>%
  rename(quantity = n)
```


### Plotting Number of Items Ordered by Aisle, where number of items is greater than 10,000
*Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.*
```{r fig.width = 8, fig.asp = 1}
aisle_df %>% 
  filter(quantity > 10000) %>% 
  ggplot(aes(x = quantity, y = reorder(aisle, quantity))) +
  geom_bar(stat = "identity") +
  labs(
    x = "Number of Items Ordered",
    y = "Aisle"
  )
```
This plot shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. The plot is in descending order based on quantity.

### Popular Items by Aisles
*Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.*
```{r}
target = c("baking ingredients", "dog food care", "packaged vegetables fruits")

instacart %>% 
  select(aisle, product_name) %>% 
  filter(aisle %in% target) %>%
  group_by(aisle, product_name) %>% 
  count(product_name) %>% 
  arrange(desc(n)) %>% 
  group_by(aisle) %>% 
  slice(1:3) %>% 
  knitr::kable()
```


### Pink Lady Apples and Coffee Ice Cream
*Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).*
```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select(product_name, order_hour_of_day, order_dow) %>% 
  arrange(order_dow) %>% 
  mutate(order_dow = recode(order_dow, `0` = "Sunday", `1` = "Monday", `2` = "Tuesday", `3` = "Wednesday", `4` = "Thursday", `5` = "Friday", `6` = "Saturday")) %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = "order_hour_of_day",
    values_fn = mean
  ) %>% 
  knitr::kable()
```


## Problem 2

First, load the `BRFSS` dataset...
```{r}
data("brfss_smart2010")
```

Then, clean the dataset...
```{r}
brfss_df = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(
    topic == "Overall Health", 
    response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")
    ) %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered = TRUE))
```


### States with 7 or more locations
*In 2002, which states were observed at 7 or more locations? What about in 2010?*
```{r}
state02 = 
  brfss_df %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  count(locationabbr) %>% 
  filter(n >= 7)

state10 =
  brfss_df %>% 
  filter(year == 2010) %>% 
  group_by(locationabbr) %>% 
  count(locationabbr) %>% 
  filter(n >= 7)
```
In **2002**, the following states were observed at 7 or more locations: `r unique(pull(state02, locationabbr))`.

And in **2010** it was these states: `r unique(pull(state10, locationabbr))`


### Excellent responses
*Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state*
```{r, fig.width = 8, fig.asp = 1}
excell_df =
  brfss_df %>% 
  filter(response == "Excellent") %>% 
  group_by(locationabbr, year) %>% 
  mutate(meandv = mean(data_value)) %>% 
  select(year, locationabbr, meandv) %>% 
  distinct(locationabbr, .keep_all = TRUE)

excell_df %>% 
  ggplot(aes(x = year, y = meandv, group = locationabbr, color = locationabbr)) +
  geom_line()
```


### Distribution of data_value for Poor to Excellent in NY State
*Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.*
```{r}
brfss_df %>% 
  filter(year %in% c(2006, 2010), locationabbr == "NY") %>% 
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_grid(. ~ year) +
  labs(
    x = "Response",
    y = "Data Value"
  )
```

The median values and interquartile range in each response group appear to be the similar across both years.  

## Problem 3

Load and tidy the accel_data.csv dataset...
```{r}
accel_df = 
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute_of_day",
    names_prefix = "activity_",
    values_to = "activity"
  ) %>% 
  mutate(
    day_type = ifelse(day %in% c("Saturday", "Sunday"), "weekend", "weekday"),
    day_type = factor(day_type),
    day = factor(day),
    minute_of_day = as.numeric(minute_of_day)
  )
```
There are `r ncol(accel_df)` variables in the `accel_df` dataframe including: `r ls(accel_df)`. Each row of the dataset includes the 63 year-old male's activity counts at each minute of a 24-hour day. The dataset includes `r nrow(accel_df)` observations.


### Total activity counts for each day
*Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?*
```{r}
agg_day_df =
  accel_df %>% 
  group_by(day_id) %>% 
  mutate(total_day_activity = sum(activity)) %>% 
  distinct(day_id, .keep_all = TRUE) %>% 
  select(-activity, -minute_of_day) %>% 
  relocate(day_type, .after = day) 

agg_day_df %>% 
  knitr::kable()
```
From this table, we see that activity tends to be lower on weekends than on weekdays.

### 24-hour activity for each day of the week
*Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.*
```{r fig.width = 8, fig.asp = 1}
accel_df %>% 
  ggplot(aes(x = minute_of_day, y = activity, color = day)) +
  geom_point(alpha = .5) + 
  geom_smooth() + 
  labs(
    x = "Minute of the 24-hour day",
    y = "Activity counts"
  )
```

From the plot, we can see that activity begins around minute 250 of the day (approx 4 am) and peaks and decreases around minute 1250 of the day (approx 8pm). The times of the day with the most activity are around minute 750 (approx 12:30pm) and between minute 1000-1250 (approx 4-8pm).