---
title: "p8105_hw3_raj2152"
author: "Rachelle Juan"
date: "10/20/2021"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)
```

## Problem 1

### Load the Instacart data for `p8105.datasets`
```{r}
data("instacart")
```

The Instacart dataset has `r nrow(instacart)` observations and `r ncol(instacart)` variables. Each row or observation corresponds to a single product within an Instacart order, therefore, one order may map to several products (i.e., there is a one-to-many relationship between orders and products). The variables provide details on each order, including things like `order_id` and `product_id` as well as the `user_id` of who placed the order. The variables also include information on where each product is located, for example: `aisle` and `department`.

* There are `r length(unique(pull(instacart, aisle_id)))` aisles. The aisles with the most items ordered from are: fresh vegetables, fresh fruits, packaged vegetables fruits, yogurt, and packaged cheese. 
```{r}
aisle_df = 
  instacart %>%
  count(aisle) %>% 
  arrange(desc(n)) %>%
  rename(quantity = n)
```


* This plot shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. The plot is in descending order based on quantity. 
```{r}
aisle_df %>% 
  filter(quantity > 10000) %>% 
  ggplot(aes(x = quantity, y = reorder(aisle, quantity))) +
  geom_bar(stat = "identity") +
  labs(
    x = "Number of Items Ordered",
    y = "Aisle"
  )
```


* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
target = c("baking ingredients", "dog food care", "packaged vegetables fruits")

instacart %>% 
  select(aisle, product_name) %>% 
  filter(aisle %in% target) %>%
  group_by(aisle, product_name) %>% 
  count(product_name) %>% 
  arrange(desc(n)) %>% 
  group_by(aisle) %>% 
  slice(1:3) %>% 
  knitr::kable()
```


* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select(product_name, order_hour_of_day, order_dow) %>% 
  arrange(order_dow) %>% 
  mutate(order_dow = recode(order_dow, `0` = "Sunday", `1` = "Monday", `2` = "Tuesday", `3` = "Wednesday", `4` = "Thursday", `5` = "Friday", `6` = "Saturday")) %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = "order_hour_of_day",
    values_fn = mean
  ) %>% 
  knitr::kable()
```


